%!/usr/bin/env pdflatex
%-*- coding: utf-8 -*-
%@author : Romain Graux
%@date : 2022 May 03, 12:22:56
%@last modified : 2022 June 06, 12:52:29

As we produce more data every year, we need to find a way to store it efficiently. Currently, we store our data in big data center consuming a lot of energy to keep these informations in electronic devices, \illustrate{Find the consumption of data centers}

It would be a good idea to find a way to store our data in a more efficient and ecological way. For storing information, hard drives don't hold a candle to DNA. Our genetic code packs billions of gigabytes into a single gram. A mere milligram of the molecule could encode the complete text of every book in the Library of Congress and have plenty of room to spare. \cite{bib:dna_data_storage}

% Since it is now possible to manipulate synthetic DNA \illustrate{Find an article talking about the first DNA synthesize}, it would be one of a new type 

But it can not be applied to all data types, for example, it is not possible yet to replace an USB stick by a DNA based USB stick and expecting the same experience. The information retrieval latency and high cost of the DNA sequencer and other instruments "currently makes this impractical for general use," says Daniel Gibson, a synthetic biologist at the J. Craig Venter Institute in Rockville, Maryland, "but the field is moving fast and the technology will soon be cheaper, faster, and smaller." Gibson led the team that created the first completely synthetic genome, which included a "watermark" of extra data encoded into the DNA. \cite{bib:dna_data_storage}

This does not mean that there are no applications for DNA based storage. DNA based storage can be used for long term media preservation archives (so called cold media storage) which are infrequently accessed and thus do not need low information retrieval latency.


\subsection{Constraints}

Unfortunately, nucleic acids have biological constraints and can not be assembled in any order like it is the case for binary digits. The DNA strands have to be created in a way that the double helix binds well together and is not immediately desctructed. We must therefore respect the biological constraints to build strong strands that can last for a certain period of time. 

In this part, we are going to go through some of the constraints that we have to respect to build a DNA strand and be able to recover it when sequencing it. Unfortunately, the list of constraints is not exhaustive and in the real life, each arangement of nucleic acids has an impact on the strength of a strand, therefore we can only simulate the longevity of a strand thanks to the actual discoveries but not strictly respect the biological constraints. 

All constraints could be reduced to limitations regarding GC content, long strands of a single nucleotide (so-called homopolymers), several repeated subsequences in a strand and motifs with biological relevance. 

\subsubsection{Synthesis}

For example, to synthesis synthetic DNA, \textit{in silico} designed constructs have to be split into smaller fragments [usually 200–3000 base pairs (bp)] \cite{bib:101038}. The fragments are then splitted into several oligonucleotides (so-called oligos) [usually 40-100 bp] that are individually synthetized. Once synthetized, all oligos are merged back together with either ligase or polymerase-based methods. One of the constraints on the GC content comes from the fact that depending on the synthetis method and the overall GC content of a fragment,  the GC content of each oligo has to be within a specific range. In oligos with high GC content, neighboring guanines tend to form an increased amount of hydrogen bonds, leading to inter/intra-strand folding \cite{bib:101371}.
To assemble oligos into larger fragments, the melting temperature (and thus the GC content) should only deviate slightly between oligos. To adhere to this restriction, the designed DNA fragments should be homogenous with respect to GC content. Homopolymers further increase the synthesis complexity, leading to fragments that are only synthesizable by using modified oligos and more sophisticated assembly methods, resulting in increased synthesis costs.

\subsubsection{PCR: Polymerase Chain Reaction}

The amplification of DNA using polymerase chain reaction (PCR) is indispensable for biological science. From DNA synthesis over cloning to DNA sequencing, PCR is used in a wide range of applications. 
One important factor of a successful PCR is the base composition of the amplification substrate. High melting temperatures due to high GC content of the DNA fragments hinder the separation of strands during the denaturation phase of the PCR. 
This reduces the yield of the PCR process, since the polymerase cannot efficiently synthesize the growing strand in the presence of previously existing hydrogen bonds. 
Stretches of repetitive DNA or high GC content can lead to the formation of secondary structures, hindering the elongation of the growing strand. 
Repetitive regions, as well as homopolymers, can also lead to polymerase slippage, a process in which polymerase briefly loses the connection to the template strand and reconnects at a different position with similar nucleotides content \cite{bib:102144}.

% \subsubsection{Storage}
% 
% \subsubsection{Sequencing}


\subsection{Requirements}

\tocontinue{}

\subsection{MESA: Mosla Error Simulator}

In order to simulate without going through an expensive and long process that is DNA synthesis and sequencing, we are going to use a simulator that takes into account a large majority of biological constraints.
This simulator has been introduced \cite{bib:10.1093/bioinformatics/btaa140} in March 2020 and is a web application for the assessment of DNA fragments in terms of guanine-cytosine (GC) content, homopolymer occurrences and length, repeating subsequences and undesirable sequence motifs.
Furthermore, MESA contains a mutation simulator, using either the error probabilities of the assessment calculation, literature-based or user-defined error rates and error spectra. MESA is fully customizable using an easy-to-use web interface, without requiring programming experience. All functionality of MESA is also contained in a REST API, enabling the incorporation of MESA evaluations into custom workflows for high-throughput evaluation and error simulation of DNA.

As we have seen in the previous section, DNA has a lot of constraints during the synthesis, storage, PCR and sequencing step; With this simulator it is now possible to have a feedback of the strength of a particular DNA strand and thus help us to move towards the best DNA coding to ensure good information retrieval in the end.

\tocontinue{}
